---
title: "Quick Demo 1: Fitting a Logistic Regression Based Generalized SuSiE (GSuSiE) Model"
author: "Ming Yuan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Demo 1: Fitting a Logistic Regression Based Generalized SuSiE (GSuSiE) Model}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 5, fig.height = 4, fig.align = "center", fig.cap = "&nbsp;", 
  dpi = 120
)
```

In this vignette, we show how to apply the `gsusie()` function on a logistic
regression based GSuSiE model dedicated to modelling binary responses.

## Get genotype data and generate binary outcome

We simulate response using genotype data from an R package named
`bigsnpr`. 
The explanatory variables `X` in this example, 
saved in `./data/genotype/genotype-subset-1.rds`, 
is a subset of genotype data containing 2000 successive SNPs of 500 individuals. 
The data are extracted using the following code: 

```{r, eval = F}
# install.packages("bigsnpr")
library(bigsnpr)

`%&%` <- function(a,b) paste0(a,b)
data.dir <- "../data/genotype/"
dir.create(data.dir, recursive=TRUE, showWarnings=FALSE)

.bed.file <- data.dir %&% "1000G_phase3_common_norel.bed"
if (!file.exists(.bed.file)) {
  bigsnpr::download_1000G(data.dir)
}

.bk.file <- data.dir %&% "1000G_phase3_common_norel.rds"
if (!file.exists(.bk.file)){
  BED <- snp_readBed(.bed.file)
}
dat <- snp_attach(.bk.file)$genotypes

nn <- 500
pp <- 2000

set.seed(12345)
startpoint <- sample(1 : (ncol(dat)-pp), size = 1)
if (nn < 2490) {ii.idx <- sample(1 : 2490, size = nn)}
X <- dat[ii.idx, startpoint : (startpoint+pp-1)]

example.data.dir <- "./example-data/"
saveRDS(X, file = example.data.dir %&% "genotype-subset-1.rds")
```


We set three variables, `X83`, `X1133`, and `X1406`, as 
the effect variables, each having a non-zero
effect size of 1 (before scaling). 
Then, we generate the linear predictor `eta`. 
We scale the linear combination to avoid unexpected large or 
small values in `eta`. 

```{r}
`%&%` <- function(a,b) paste0(a,b)
example.data.dir <- "./example-data/"

X <- readRDS(example.data.dir %&% "genotype-subset-1.rds")
nn <- nrow(X)  # 500
pp <- ncol(X)  # 2000

n_effect_vars <- 3
## effect variables with non-zero effects
effect_idx <- c(83, 1133, 1406)

## effect size (before scaling)
effect_size <- rep(1, times = n_effect_vars)

## linear predictor
eta <- scale(X[,effect_idx, drop=F] %*% as.matrix(effect_size))

## response
expit <- function(eta) {
  ifelse(eta > 0, 1 / (1 + exp(-eta)), exp(eta) / (1 + exp(eta)))
}
set.seed(20240220)
y <- rbinom(nn, 1, expit(eta))
```

## Fit a logistic regression based GSuSiE model

When fitting a GLM-based GSuSiE model, we prefer adding an all-one column to 
introduce an intercept term. In practice, we are inclined to add this auxiliary 
column to the end of the design matrix X such that variable indices would not 
be confusing when printing the summarized results by the `summary()` function. 
This step is optional; nevertheless, we find that, by adding this all-one 
column, the model is less likely to be overfitted. (Wooh!)

By default, we assume there are no more than 10 effect variables
and set `maxL = min(10, ncol(X))`. You may choose to specify another
value to `maxL`.

To specify fitting a **logistic regression based GSuSiE model**, 
use the argument `family = "binomial"`. Correspondingly, the
`gsusie()` function call is:

```{r}
library(gsusie)
res_gs <- gsusie(cbind(X, 1), y, family = "binomial")
```

### Posterior inclusion probabilities

The `print_gsusie_coefficients()` function offers 10 (by default) variables
with the highest posterior inclusion probabilities (PIPs). By default,
we output the 95% credible intervals of the posterior estimation of the coefficients. 

```{r}
print_gsusie_coefficients(res_gs)
```

### Credible sets (CS)

As with the SuSiE method, we output 95% credible sets by default.

```{r}
print(res_gs$sets)
```

The three relevant signals have been captured by the 3 credible sets
here. Effect variables `X83` and `X1406` are detected, 
each discovered along with another highly-correlated null variable. 

### Graphical display of PIPs and credible sets

The `gsusie_plot()` function provides plots for the PIP of each variable.

By default, since the intercept is just an auxiliary variable, we do not
include it in the plot. In this case, the argument `intercept_index` needs
to be specified. (P.S. By default, we set `include_intercept = FALSE` to
remove the intercept term on the plot.)

```{r, fig.height=3, fig.width=5}
gsusie_plot(res_gs, y = "PIP", 
            include_intercept = FALSE, 
            intercept_index = (pp+1), 
            effect_indices = effect_idx)
```

By specifying `effect_indices`, the true effect variables are
colored in red. The 95% credible sets are identified as circled in
different colors. 

## Some additional words

As you may notice, the `gsusie()` function provides an option for robust estimation. 
Well... yes. The robust approaches in this package aim to down-weight the impact 
of outliers, i.e., the extremely big response values, in the model fitting. 
These approaches are effective in fitting a Poisson regression based 
GSuSiE model. 
However, they do not demonstrate a comparable advantage over a logistic 
regression based GSuSiE model; 
what's worse, those robust approaches could be harmful. 
That is because, unlike the count data in which some big numbers may occur,
the distribution of a binary response does not spread out; 
the value of a binary response is just one or zero. 
As a result, robust estimation is unnecessary. 

